---
title: "Forest plots for Handley et al. 2021 & 2022 length of stay studies"
author: "Heather H. Burris (with Nancy Yang and Anne Mullin)"
date: '2021-11-17'
output: 
  html_document:
    theme: flatly
    toc: true
    toc_float:
      collapsed: false
    code_folding: hide
    code_download: true
---

Associated papers:

- Handley et al. 2021: [Birth Hospital Length of Stay and Rehospitalization During COVID-19 | Pediatrics | American Academy of Pediatrics](https://publications-aap-org.proxy.library.upenn.edu/pediatrics/article/149/1/e2021053498/183458/Birth-Hospital-Length-of-Stay-and)
- Handley et al. 2022: [Postpartum Length of Stay and Hospital Readmission Before and During the Coronavirus Disease 2019 (COVID-19) Pandemic: Obstetrics & Gynecology](https://journals.lww.com/greenjournal/Fulltext/2022/03000/Postpartum_Length_of_Stay_and_Hospital_Readmission.5.aspx)

The original R code written by Heather Burris was adapted for this R Markdown file by Silvia Canelón.

The code chunks are hidden by default, but you can Show or Hide all code using the "Code" dropdown menu on the top right of this page. Or you can selectively view specific code chunks by clicking the "Code" buttons througout the page.

You can download this R Markdown file (.Rmd) by using the "Code" dropdown menu. You can also find the R Markdown file along with the original R script, data, and plot files organized in the Burris Lab Data Viz Resources folder. Each is organized according to the directory tree below.

- R Markdown file and the corresponding HTML file are found at the root of the folder "handley-LOS"
- Code is in the "code" folder
- Data are in the "data" folder"
- Plot files are in the "output" folder.

```
.
├── Burris-Lab/data-viz-resources
└── forest-plot
    └── handley-LOS
        ├── README.md
        ├── forest-plot-handley-LOS.Rmd
        ├── forest-plot-handley-LOS.html
        ├── code
        ├── data
        └── output

```

Detailed directory tree:

```
.
├── Burris-Lab/data-viz-resources
└── forest-plot
    └── handley-LOS
        ├── README.md
        ├── forest-plot-handley-LOS.Rmd
        ├── forest-plot-handley-LOS.html
        ├── code
        │   └── Maternal paper figures r code20211117.R
        ├── data
        │   ├── Data for Figure of mat readmit1121.csv
        │   └── dataforLMatLOS1121.csv
        └── output
            ├── PlotsSara_cosmos_fig_LOSONLY2noannote1.tiff
            ├── PlotsSara_cosmos_fig_MATERNALreadmitONLY2noannote1.tiff
            ├── bPlotsSara_cosmos_fig_MATERNALLOSONLY3withEE.tiff
            ├── cPlotsSara_cosmos_fig_MATERNALBOTH3withee.tiff
            └── cPlotsSara_cosmos_fig_MATERNALreadmitONLY3withee.tiff
```

## Setup

```{r setup, include=TRUE, message=FALSE, warning=FALSE}
xfun::pkg_attach("table1",
                 "ggplot2",
                 "dplyr",
                 "scales",
                 "grid",
                 "gridExtra",
                 "lattice",
                 "gtable",
                 "cowplot",
                 "here")

knitr::opts_chunk$set(echo = TRUE)

# define directory

# if you opened the R project file burris-lab-data-viz.Rproj:
dir <- here("forest-plot", "handley-LOS")

# if you're setting the working directory manually:

# getwd() to get the current working directory
# example output: "C:/Users/hhers/Dropbox"

# setwd() to set your working directory
# example: 
# dir <- setwd("C:/Users/hhers/Dropbox/Burris Lab Data Viz Resources/forest-plot/handley-LOS")
```

## Short LOS

```{r import-data-losmat}
LOSMAT.Fig <- read.csv(here(dir, "data", "dataforLMatLOS1121.csv"))

LOSMAT.Fig
#View(LOSMAT.Fig)
```

```{r}
df1 <- data.frame(LOSMAT.Fig)
df1
df1$combined <- factor(
  df1$combined, 
  levels=rev(df1$combined) # flips the order of the factors
  ) 
#df1 %>% 
#  mutate(stripe = factor(ifelse(DOY %% 2 == 0, 1, 0))) %>% 

# p1 ----
p1 <- ggplot(df1, aes(x=combined, 
                      y=OR, 
                      ymin=lower, 
                      ymax=upper)) + 
  geom_linerange(size=1, 
                 col="black") +
  geom_hline(yintercept=1, 
             lty=2) +
  geom_point(size=1.5, 
             colour="black", 
             stroke = 1.0, 
             shape=16) +
  geom_rect(data=df1, 
            colour="white",
            mapping=aes(
              xmin=as.numeric(combined)-0.5,
              xmax=as.numeric(combined)+0.5, 
              ymin=0.3, 
              ymax=3.6, 
              fill=Model),
            alpha = 0.5)+
  # scale_fill_manual( = alpha(c("light blue", "red"), 0.2))+
  labs(
    y="Odds ratios of short postpartum hospitalization LOS",
    x="Mode of delivery")+
  scale_y_continuous(trans = "log10", 
                     limits=c(0.3, 7), 
                     n.breaks = 9) +
  theme_classic()+
  theme(
    #legend.title = element_blank(),
    #legend.text=element_text(size=11),
    legend.position = "none",
    axis.text.x = element_text(color = "black", 
                               size =  10),
    axis.text.y = element_blank(),
    axis.title.x = element_text(color = "black", 
                                size =  18),
    axis.title.y = element_text(color = "black", 
                                size =  18),
    axis.line.y = element_blank(),
    axis.ticks.y = element_blank())+
  #axis.text.y = element_text(color = "black", size =  10)) +
  coord_flip() 

p1

#p1<-p1+scale_fill_brewer()
#p1<-p1+scale_fill_manual(values = alpha(c("light blue", "red"), .1))

# p1a ----
p1a<-p1+ geom_rect(data=df1, colour="white", mapping=aes(xmin= 0.5, xmax=2.5, ymin=0.3, ymax=0.4), fill="dark green", alpha =0.5)
#names(df1)

# p1b ----
p1b<-p1a+geom_rect(data=df1, colour="white", mapping=aes(xmin= 2.5, xmax=4.5, ymin=0.3, ymax=0.4), fill="dark green", alpha =0.5)

p1b

# p1c ----
p1c <- p1b + geom_rect(data=df1, 
                       colour="white", 
                       mapping=aes(xmin= 4.5,
                                   xmax=6.5, 
                                   ymin=0.3, 
                                   ymax=0.4), 
                       fill="dark green", 
                       alpha =0.5)

p1c
p1c <- p1c + scale_fill_brewer(palette = "Greens")
p1c
p1c <- p1c + annotate(geom="text", 
                      x=6.0,
                      y=3.0,
                      label="A",
                      size=5)
p1c <- p1c + annotate(geom="text", 
                      x=6,  
                      y=0.5, 
                      label="Unadjusted",
                      size=4)
p1c <- p1c + annotate(geom="text",
                      x=5,
                      y=0.5, 
                      label="Adjusted",
                      size=4)
p1c <- p1c + annotate(geom="text", 
                      x=4,
                      y=0.5,
                      label="Unadjusted",
                      size=4)
p1c <- p1c + annotate(geom="text",
                      x=3,
                      y=0.5,
                      label="Adjusted",
                      size=4)
p1c <- p1c + annotate(geom="text",
                      x=2,
                      y=0.5,
                      label="Unadjusted",
                      size=4)
p1c <- p1c + annotate(geom="text",
                      x=1,
                      y=0.5,
                      label="Adjusted",
                      size=4)
p1c

# p1d ----
p1d <- p1c + annotate(geom="text", 
                      x=5.5,
                      y=0.35,
                      label="All",
                      size=5,
                      colour="white",
                      angle=90)
p1d <- p1d+annotate(geom="text",
                    x=3.5,
                    y=0.35,
                    label="Vaginal",
                    colour="white", 
                    size=5,
                    angle=90)
p1d <- p1d+annotate(geom="text", 
                    x=1.5,
                    y=0.35, 
                    label="Cesarean",
                    colour="white",
                    size=5,
                    angle=90)
p1d
p1d <- p1d + theme(plot.margin=unit(c(1,1,1,1),
                                    "cm")) #adding space to the bottom of the first grob (c(bottom,left,top,right))
p1d
```

### Maternal LOS, only 3 with EE

```{r}
## p1e ----
p1e <- p1d + annotate(geom="text", 
                      x=6,
                      y=5,
                      label="1.99 (1. 97 -2.02)",
                      colour="black", 
                      size=3)
p1e <- p1e + annotate(geom="text", 
                      x=5,  
                      y=5, 
                      label="2.35 (2. 32 -2. 3 9)",
                      colour="black", 
                      size=3)
p1e <- p1e + annotate(geom="text", 
                      x=4,
                      y=5,
                      label="1.91 (1. 88 -1. 9 4)",
                      colour="black", 
                      size=3)
p1e <- p1e + annotate(geom="text",
                      x=3,
                      y=5,
                      label="2.14 (2. 11 -2. 1 8)",
                      colour="black",
                      size=3)
p1e <- p1e + annotate(geom="text",
                      x=2,
                      y=5,
                      label="2.25 (2. 21 -2. 3 0)",
                      colour="black",
                      size=3)
p1e <- p1e + annotate(geom="text",
                      x=1,
                      y=5,
                      label="2.90 (2.83-2.98)",
                      colour="black",
                      size=3)
p1e

ggsave(here(dir, "output", "bPlotsSara_cosmos_fig_MATERNALLOSONLY3withEE.tiff"),
       width=7.2,
       height=4.8,
       units= "in",
       device = "tiff",
       dpi=400,
       plot = p1e)
```

## Readmission figure

```{r import-data-readmit}
readmitMAT.Fig <- read.csv(here(dir, "data", "Data for Figure of mat readmit1121.csv"))

readmitMAT.Fig
# View(readmitMAT.Fig)
```

```{r}
df2 <- data.frame(readmitMAT.Fig)
df2

df2$combined <- factor(df2$combined, levels=rev(df2$combined)) # flips the order of the factors
#df2 %>% 
#  mutate(stripe = factor(ifelse(DOY %% 2 == 0, 1, 0))) %>% 
```

```{r}
# p2 ----
p2 <- ggplot(df2, aes(x=combined, 
                      y=OR, 
                      ymin=lower, 
                      ymax=upper)) + 
  geom_linerange(size=1, 
                 col="black") +
  geom_hline(yintercept=1, 
             lty=2) +
  geom_point(size=1.5, 
             colour="black", 
             stroke = 1.0, 
             shape=16) +
  geom_rect(data=df2, 
            colour="white", 
            mapping=aes(xmin=as.numeric(combined)-0.5, 
                        xmax=as.numeric(combined)+0.5, 
                        ymin=0.3, 
                        ymax=3.6, 
                        fill=Model),
            alpha = 0.5)+
  # scale_fill_manual( = alpha(c("light blue", "red"), 0.2))+
  labs(y="Odds ratios of rehospitalization",
       x="Mode of delivery")+
  scale_y_continuous(trans = "log10", 
                     limits=c(0.3, 7), 
                     n.breaks = 9) +
  theme_classic()+
  theme(
    #legend.title = element_blank(),
    #legend.text=element_text(size=11),
    legend.position = "none",
    axis.text.x = element_text(color = "black", 
                               size =  10),
    axis.text.y = element_blank(),
    axis.title.x = element_text(color = "black", 
                                size =  18),
    axis.title.y = element_text(color = "black", 
                                size =  18),
    axis.line.y = element_blank(),
    axis.ticks.y = element_blank())+
  #axis.text.y = element_text(color = "black", size =  10)) +
  coord_flip() 
#p2<-p2+scale_fill_brewer()
#p2<-p2+scale_fill_manual(values = alpha(c("light blue", "red"), .1))
p2

# p2a ----
p2a<-p2+ geom_rect(data=df2, 
                   colour="white", 
                   mapping=aes(xmin= 0.5, 
                               xmax=2.5, 
                               ymin=0.3, 
                               ymax=0.4), 
                   fill="dark green", 
                   alpha =0.5)
#names(df2)

# p2b ----
p2b<-p2a+geom_rect(data=df2, 
                   colour="white", 
                   mapping=aes(xmin= 2.5, 
                               xmax=4.5, 
                               ymin=0.3, 
                               ymax=0.4), 
                   fill="dark green", 
                   alpha =0.5)

p2b

# p2c ----
p2c <- p2b+geom_rect(data=df2, 
                     colour="white", 
                     mapping=aes(xmin= 4.5, 
                                 xmax=6.5, 
                                 ymin=0.3, 
                                 ymax=0.4), 
                     fill="dark green", 
                     alpha =0.5)

p2c
p2c <- p2c+scale_fill_brewer(palette = "Greens")
p2c
p2c <- p2c + annotate(geom="text", 
                      x=6.0,  
                      y=3.0, 
                      label="B", 
                      size=5)
p2c <- p2c+ annotate(geom="text", 
                     x=6,  
                     y=0.5, 
                     label="Unadjusted", 
                     size=4)
p2c <- p2c+ annotate(geom="text", 
                     x=5,  
                     y=0.5, 
                     label="Adjusted", 
                     size=4)
p2c <- p2c+ annotate(geom="text", 
                     x=4,
                     y=0.5,
                     label="Unadjusted",
                     size=4)
p2c <- p2c+ annotate(geom="text",
                     x=3,
                     y=0.5,
                     label="Adjusted",
                     size=4)
p2c <- p2c+ annotate(geom="text",
                     x=2,
                     y=0.5,
                     label="Unadjusted",
                     size=4)
p2c <- p2c+ annotate(geom="text",
                     x=1,
                     y=0.5,
                     label="Adjusted",
                     size=4)
p2c

# p2d ----
p2d <- p2c + annotate(geom="text", 
                      x=5.5,
                      y=0.35, 
                      label="All", 
                      size=5, colour="white", 
                      angle=90)
p2d <- p2d + annotate(geom="text", 
                      x=3.5, 
                      y=0.35,
                      label="Vaginal",
                      colour="white",
                      size=5,
                      angle=90)
p2d <- p2d + annotate(geom="text",
                      x=1.5,
                      y=0.35, 
                      label="Cesarean",
                      colour="white", 
                      size=5,
                      angle=90)
p2d
p2d = p2d + theme(plot.margin=unit(c(1,1,1,1),"cm")) #adding space to the bottom of the first grob (c(bottom,left,top,right))
p2d
```

### Maternal readmit, only 3 with EE

```{r}
# p2e ----
p2e <- p2d + annotate(geom="text", 
                      x=6,
                      y=5,
                      label="1.08 (1.03-1.13)",
                      colour="black",
                      size=3)
p2e <- p2e + annotate(geom="text", 
                      x=5,
                      y=5,
                      label="1.02 (0.97-1.08)",
                      colour="black",
                      size=3)
p2e <- p2e + annotate(geom="text",
                      x=4,
                      y=5,
                      label="1.07 (1.00-1.15)",
                      colour="black", 
                      size=3)
p2e <- p2e + annotate(geom="text", 
                      x=3,
                      y=5,
                      label="1.03 (0.96-1.10)",
                      colour="black", 
                      size=3)
p2e <- p2e + annotate(geom="text",
                      x=2,
                      y=5,
                      label="1.10 (1.02-1.18)",
                      colour="black",
                      size=3)
p2e <- p2e + annotate(geom="text",
                      x=1,
                      y=5,
                      label="1.01 (0.94-1.08)",
                      colour="black",
                      size=3)
p2e

ggsave(here(dir, "output", "cPlotsSara_cosmos_fig_MATERNALreadmitONLY3withee.tiff"),
       width=7.2,
       height=4.8,
       units= "in",
       device = "tiff",
       dpi=400,
       plot = p2e)

```

### Maternal, both 3 with EE

```{r}
# p1e + p2e ----

# arranging the grobs
both <-grid.arrange(
  p1e,
  p2e,
  ncol = 1,
  nrow = 2,
  heights = c(4, 4),
  clip = FALSE
)
both

#saving as tiff (can save as any file type here)
ggsave(here(dir, "output", "cPlotsSara_cosmos_fig_MATERNALBOTH3withee.tiff"),
       width=7.2,
       height=9.6,
       units= "in",
       device = "tiff",
       dpi=400,
       plot = both)
getwd()
```

## No annotation

### Maternal LOS, only 2

```{r}
# p1c ----
p1c <- p1b + geom_rect(data=df1,
                       colour="white",
                       mapping=aes(xmin= 4.5, 
                                   xmax=6.5, 
                                   ymin=0.3, 
                                   ymax=0.4), 
                       fill="dark green", 
                       alpha =0.5)

p1c
p1c <- p1c + scale_fill_brewer(palette = "Greens")
p1c
#p1c <- p1c + annotate(geom="text", x=6.0,  y=3.0, label="A", size=5)
p1c <- p1c+ annotate(geom="text", 
                     x=6,
                     y=0.5,
                     label="Unadjusted",
                     size=4)
p1c <- p1c+ annotate(geom="text",
                     x=5,
                     y=0.5,
                     label="Adjusted",
                     size=4)
p1c <- p1c+ annotate(geom="text",
                     x=4,
                     y=0.5,
                     label="Unadjusted",
                     size=4)
p1c <- p1c+ annotate(geom="text",
                     x=3,
                     y=0.5,
                     label="Adjusted",
                     size=4)
p1c<-p1c+ annotate(geom="text",
                   x=2,
                   y=0.5,
                   label="Unadjusted",
                   size=4)
p1c<- p1c+ annotate(geom="text",
                    x=1,
                    y=0.5,
                    label="Adjusted",
                    size=4)
p1c

# p1d ----
p1d<-p1c+annotate(geom="text",
                  x=5.5,
                  y=0.35, 
                  label="All",
                  size=5,
                  colour="white",
                  angle=90)
p1d<-p1d+annotate(geom="text",
                  x=3.5, 
                  y=0.35,
                  label="Vaginal",
                  colour="white",
                  size=5,
                  angle=90)
p1d<-p1d+annotate(geom="text",
                  x=1.5,
                  y=0.35, 
                  label="Cesarean",
                  colour="white",
                  size=5,
                  angle=90)
p1d
p1d <- p1d + theme(plot.margin=unit(c(1,1,1,1),"cm")) #adding space to the bottom of the first grob (c(bottom,left,top,right))

ggsave(here(dir, "output", "PlotsSara_cosmos_fig_LOSONLY2noannote1.tiff"),
       width=7.2,
       height=4.8,
       units= "in",
       device = "tiff",
       dpi=400,
       plot = p1d)
```

```{r}
# p2c ----
p2b

p2c <- p2b + geom_rect(data=df2,
                       colour="white",
                       mapping=aes(xmin= 4.5, 
                                   xmax=6.5,
                                   ymin=0.3,
                                   ymax=0.4),
                       fill="dark green",
                       alpha =0.5)

p2c
p2c <- p2c + scale_fill_brewer(palette = "Greens")
p2c
#p2c <- p2c + annotate(geom="text", x=6.0,  y=3.0, label="B", size=5)
p2c <- p2c + annotate(geom="text", 
                      x=6,
                      y=0.5,
                      label="Unadjusted",
                      size=4)
p2c <- p2c + annotate(geom="text",
                      x=5,
                      y=0.5,
                      label="Adjusted",
                      size=4)
p2c <- p2c + annotate(geom="text",
                      x=4,
                      y=0.5,
                      label="Unadjusted",
                      size=4)
p2c <- p2c + annotate(geom="text",
                      x=3,
                      y=0.5,
                      label="Adjusted",
                      size=4)
p2c <- p2c + annotate(geom="text",
                      x=2,
                      y=0.5,
                      label="Unadjusted",
                      size=4)
p2c <- p2c + annotate(geom="text",
                      x=1,
                      y=0.5,
                      label="Adjusted",
                      size=4)
p2c
```

### Maternal readmit, only 2

```{r}

# p2d ----
p2d<-p2c+annotate(geom="text",
                  x=5.5,
                  y=0.35,
                  label="All",
                  size=5,
                  colour="white",
                  angle=90)
p2d<-p2d+annotate(geom="text",
                  x=3.5,
                  y=0.35, 
                  label="Vaginal",
                  colour="white", 
                  size=5,
                  angle=90)
p2d<-p2d+annotate(geom="text", 
                  x=1.5,
                  y=0.35, label="Cesarean",
                  colour="white",
                  size=5,
                  angle=90)
p2d
p2d = p2d + theme(plot.margin=unit(c(1,1,1,1),"cm")) #adding space to the bottom of the first grob (c(bottom,left,top,right))

ggsave(here(dir, "output", "PlotsSara_cosmos_fig_MATERNALreadmitONLY2noannote1.tiff"),
       width=7.2,
       height=4.8,
       units= "in",
       device = "tiff",
       dpi=400,
       plot = p2d)

# getwd()
```
